A Robust Bayesian Approach for Causal Inference Problems

### 2023-05-18 JASA-template.tex
update till simulation

\subsection{Robust Bayesian Analysis}
We perform our
robust Bayesian analysis on $q\coloneqq(q_1$, \dots, $q_p)\in\mathcal{P}$, where
\begin{equation}
	\mathcal{P} \coloneqq \mathcal{P}_1\times\cdots\times\mathcal{P}_p\subseteq \left(0, 1\right)^{p}.
\end{equation}

>>>>

\subsection{Robust Bayesian Analysis}
The hierarchical model presented above is fairly common in
Bayesian analysis and performs well when we have sufficient data to begin with. However, especially
in the case of causal inference having sufficient data may not be practical. Moreover, since there’s
an aspect of intervention in causal effect estimator, we also need to be cautious around the side
effects of a treatment.
Therefore, we are interested in having a more robust framework for selecting a variable. So
that, once we are initiating our treatment, we can decide whether we require more data to conduct
our analysis. For instance, we can assign a cost for false positives and false negatives and also for
conducting more tests. This way we can have a more generalised utility based framework with
which we can incorporate the notion of indeterminacy.
We perform our
robust Bayesian analysis on $q\coloneqq(q_1$, \dots, $q_p)\in\mathcal{P}$, where
\begin{equation}
	\mathcal{P} \coloneqq \mathcal{P}_1\times\cdots\times\mathcal{P}_p\subseteq \left(0, 1\right)^{p}.
\end{equation}

### 2023-05-18 basu23c.tex
update till simulation

\subsection{Robust Bayesian Analysis}
We perform our
robust Bayesian analysis on $q\coloneqq(q_1$, \dots, $q_p)\in\mathcal{P}$, where
\begin{equation}
	\mathcal{P} \coloneqq \mathcal{P}_1\times\cdots\times\mathcal{P}_p\subseteq \left(0, 1\right)^{p}.
\end{equation}

\subsection{Variable selection and coefficient adjustment}
For the co-variate selection, we look into the posterior expectation of $\pi_j$. 
We consider the $j$-th predictor to be removed from both the
treatment and outcome model, if
\begin{align}
	\lexp (\pi_j\mid W)\coloneqq \inf_{q_j\in \mathcal{P}_j} \text{E}(\pi_j\mid W) > 1/2.
\end{align}

For the rest of the variables, some of them will be present
in the model as confounders and some will only be associated with
either the treatment. Let $\mathcal{S}$ denote the set of
predictors such that,
\begin{equation}
	\mathcal{S}\coloneqq
	\left\{j : \lexp(\pi_j\mid W) < 1/2\right\}.
\end{equation}
That is the set that contains all the variables which are not
removed from both treatment and outcome model. Now, for
each fixed value of $q$,
let $\hat{\beta}_{\mathcal{S}}(q_{\mathcal{S}})$ be the posterior means of the regression coefficients of the outcome model with respect to
the predictors that belong to $\mathcal{S}$. Similarly,
$\hat{\gamma}_{\mathcal{S}}(q_{\mathcal{S}})$ be the posterior means of the regression
coefficients for the treatment effects. Since, we use a continuous type
selection prior, these regression coefficients are non-zero in nature.
Therefore, to adjust the sparsity, we apply the 
``decoupled shrinkage and selection'' method proposed by \cite{hahn2015}. To do so, we solve the following adaptive LASSO-type \cite{Zou2006}
problems

\begin{align}
	\hat{\beta}^*_{\mathcal{S}}(q) &= 
	\arg\min_{\beta_{\mathcal{S}}} \frac{1}{n}\|X_{\mathcal{S}}\hat{\beta}_{\mathcal{S}}(q)
	- X_{\mathcal{S}} \beta_{\mathcal{S}}\|_2^2 + \lambda\sum_{j\in\mathcal{S}} 
	\frac{|\beta_j|}{|\hat{\beta}_j(q_j)|}
\end{align}
and
\begin{align}
	\hat{\gamma}^*_{\mathcal{S}}(q) &= 
	\arg\min_{\gamma_{\mathcal{S}}} \frac{1}{n}\|X_{\mathcal{S}}\hat{\gamma}_{\mathcal{S}}(q)
	- X_{\mathcal{S}} \gamma_{\mathcal{S}}\|_2^2 + \lambda\sum_{j\in\mathcal{S}} 
	\frac{|\gamma_j|}{|\hat{\gamma}_j(q_j)|}
\end{align}
where $q_{j}\in \mathcal{P}_{j}$ for all $j\in\mathcal{S}$.

\iftrue
\subsection{Causal effect adjustment} The DSS method explained
above give us adjusted coefficient estimates for the treatment
and outcome model. However, as result the causal effect estimate
remains the same and modelling with such adjusted sparse effect 
will contribute to the prediction error. Therefore, we need to 
adjust the causal effect estimate as well. We do that in three steps
similar to double residual regression procedure as suggested by
\cite{robinson1988}.

>>>>

\subsection{Robust Bayesian Analysis}

The hierarchical model presented above is a standard spike and slab model for
variable selection and performs well when we have sufficient data to begin with. 
However, especially in the case of causal inference having sufficient data may not 
be practical. Moreover, since there’s an aspect of intervention in causal effect 
estimator, we also need to be cautious about the side effects of a treatment.
Therefore, we are particularly interested in constructing a robust Bayesian framework
for variable selection. So that, once we are preparing a guideline for treatment, we 
can decide whether we require more data to conduct our analysis. For instance, 
we want to evaluate the problem based on the cost associated with different aspects
of variable selection including the possibility of abstention that is
conducting more tests. This way we can have a generalised utility based framework with
which we can interpret the lack of evidence efficiently without having an adverse
effect.

In general, a possible outcome of mistreating a subject can have an outcome
which can not be associated with a loss function for obvious reasons. Instead, we
assume that we can always treat a subject to normality after an initial mistreatment.
and we can associate a loss function with the cost of further treatments. This way, 
for in simplest case, can associate two constant values for false positives and
false negatives. Clearly, false positives will lead to unwanted side effects and
false negatives will lead to mistreatment of the patient. Finally, we associate
a loss value for abstention which can be interpreted as the cost of further tests.
We present this notion in the following table.
\begin{table}[h]
    \centering
    \begin{tabular}{l|ccc}
       &  False Positive & False Negative & Abstention \\
       \hline
       Value  &  $\ell_1$ & $\ell_2$ & $\ell_3$
    \end{tabular}
    \caption{Losses with respect to different types of variable selection}
    \label{tab:loss:vs}
\end{table}
Ideally, in most cases, $\ell_3\ll \ell_1,\ell_2$. However, in certain scenarios,
this might not be the case. Especially when the condition of a subject deteriorates rapidly
over time.

Now, based on this notion of abstaining from selecting a variable, we can perform
a sensitivity analysis over a set of priors on the prior selection probability.
That is, we can consider a set of possible values for $q_j$ such that
$q\coloneqq(q_1$, \dots, $q_p)\in\mathcal{P}$, where
\begin{equation}
	\mathcal{P} \coloneqq \mathcal{P}_1\times\cdots\times\mathcal{P}_p\subseteq \left(0, 1\right)^{p}.
\end{equation}
Here the equality occurs for the near vacuous case. However, in real-life
situations, performing a robust Bayesian analysis for near vacuous case is 
not practical. Instead, we incorporate expert elicitation to define our model.
For instance, for the equiprobable case, we can consider $q = q_1=q_2=\cdots = q_p$
and $q\in \left[\underline{q}, \overline{q}\right]$
where $p\underline{q}$ and $p\overline{q}$ represent the bounds of the prior expectation on the
total number of variables present in either of the models.

\subsection{Variable selection and coefficient adjustment}
For the co-variate selection, we look into the posterior expectation of $\pi_j$. 
We consider the $j$-th predictor to be removed from both the
treatment and outcome model, if
\begin{align}
	\uexp (\pi_j\mid W)\coloneqq \sup_{q_j\in \mathcal{P}_j} \text{E}(\pi_j\mid W) < 1/2.
\end{align}
Similarly, we consider the $j$-th predictor to be present in at least one of the models, if
\begin{align}\label{eq:vs:sel}
	\lexp (\pi_j\mid W)\coloneqq \sup_{q_j\in \mathcal{P}_j} \text{E}(\pi_j\mid W) > 1/2.
\end{align}
Otherwise, we consider the variable to be indeterminate and abstain from putting
it in any of the models and reports lack of information.

In general, this framework is self sufficient for variable selection. However, for
model fitting and prediction, we need to evaluate the values 
of the regression coefficients. For that we first need to find the set of active
predictors with respect to our prior expectation of the selection probability $q$.
We define the set $S(q)$ such that for any fixed $q$.
\begin{equation}
	S(q)\coloneqq
	\left\{j : E(\pi_j\mid W) \ge 1/2\right\}.
\end{equation}
That is the set that contains all the variables which are active
either in the treatment model and or in the outcome model for a fixed value of $q$. 
Clearly, if we take intersection over all $q$ then it gives us back the set of
active variables obtained through \cref{eq:vs:sel}. That is 
\begin{align}
    \mathcal{S}\coloneqq \left\{j:\lexp (\pi_j\mid W)>1/2\right\}
    = \bigcap_{q\in \mathcal{P}}S(q)
\end{align}

Now, for each fixed value of $q$, let $\hat{\beta}_{S(q)}$ be the posterior means 
of the regression coefficients of the outcome model with respect to
the predictors that belong to $S(q)$. Similarly,
$\hat{\gamma}_{S(q)}$ be the posterior means of the regression
coefficients for the treatment effects. Since, we use continuous 
spike and slab priors, these regression coefficients are not sparse.
Moreover, with our variable selection we only determine whether the variable 
is included in at least one of the models. But, we can not determine a specific 
association. Therefore, to adjust the sparsity of the estimates and understand
the specific association with the treatment/outcome/both, we apply the 
``decoupled shrinkage and selection'' method proposed by \cite{hahn2015}. 
For that, we solve the following adaptive LASSO-type \cite{Zou2006}
problems

\begin{align}
	\hat{\beta}^D_{S(q)} &= 
	\arg\min_{\beta_{S(q)}} \frac{1}{n}\|X_{S(q)}\hat{\beta}_{S(q)}
	- X_{S(q)} \beta_{S(q)}\|_2^2 + \lambda\sum_{j\in S(q)} 
	\frac{|\beta_{j,S(q)}|}{|\hat{\beta}_{j,S(q)}|}
\end{align}
and
\begin{align}
	\hat{\gamma}^D_{S(q)} &= 
	\arg\min_{\gamma_{S(q)}} \frac{1}{n}\|X_{S(q)}\hat{\gamma}_{S(q)}
	- X_{S(q)} \gamma_{S(q)}\|_2^2 + \lambda\sum_{j\in S(q)} 
	\frac{|\gamma_{j,S(q)}|}{|\hat{\gamma}_{j,S(q)}|}
\end{align}
where $q\in \mathcal{P}$.

The DSS method explained
above give us adjusted coefficient estimates for the treatment
and outcome model. However, as result the causal effect estimate
remains the same and modelling with such adjusted sparse effect 
will contribute to the prediction error. Therefore, we need to 
adjust the causal effect estimate as well. Let
\begin{equation}
    \hat{T}^D = \mathbb{I}\left(\hat{\gamma}_0+X_{S(q)}\hat{\gamma}^D_{S(q)} >0\right)
    \quad\text{and}\quad
    \hat{T} = \mathbb{I}\left(\hat{\gamma}_0+X\hat{\gamma}^q >0\right).
\end{equation}
and let $\beta_{T}^D(q)$
be the causal effect after applying `DSS' and $\hat{\beta}_{T}(q)$ be the
posterior estimate of the causal effect for any fixed $q\in\mathcal{P}$. Then
the adjusted estimate is given by:
\begin{equation}
    \beta_{T}^D(q) = \left(\left(\hat{T}^D\right)^T \hat{T}^D\right)^{-1}
    \left(\hat{T}^D\right)^T \left[\hat{\beta}_{T}(q)\hat{T} 
    + X_{S(q)}\left(\hat{\beta}_{S(q)} - \hat{\beta}^D_{S(q)}\right) 
    +X_{S^C(q)}\hat{\beta}_{S^C(q)}\right]
\end{equation}
where
\begin{equation}
    S^C(q)\coloneqq \left\{1,2,\cdots, p\right\} \neg S(q).
\end{equation}

\iffalse
\subsection{Causal effect adjustment}  We do that in three steps
similar to double residual regression procedure as suggested by
\cite{robinson1988}.
